Machine Learning guided workflow for Ribosome Binding Site engineering

### 2021-04-08 paper/main.tex
Added characteristics figure.

A genetic part library should display a number of characteristics to be deemed useful.
First quality expected from a good library is a wide range of relevant trait values (in our case it is the TIR) so that relevant genetic devices can be appropriately optimised.
Second important feature of a library is the diversity of offered sequences, which ensures that sequences compatible with different assembly methods and plasmids are available. 
Additionally, it is very beneficial for the library to show a number of parts with similar numerical values for relevant trait, but with different sequences, so that a designer would have to settle on suboptimal part due to, for example, assembly incompatibility.
Finally, a good library has to be reliable - if the numbers provided in the library are unreliable designers will be reluctant to use.
Figure X shows how library conforms to these requirements.
In short, due to focusing on exploitation distribution of our TIRs is skewed toward strong ones, but there is still a significant amount of RBSs with medium and low strengths.
High diversity of our sequences is shown by lack of relative lack of bias in position weight matrices for different bins.
Finally, we were able to keep the average standard deviation for each bin within XX\% with 6 biological replicates for each sample, which we find satisfactory for reliability.\\
Our library is also showing two characteristics that may be useful in future RBS design efforts.
One of them is the perceived editing distance between two sequences required for meaningful improvement in the TIR. 
We define the editing distance as Hamming distance, that is, how many positions have to be changed to get from one sequence to the other (Hamming distance of 0 means that the sequences are identical and 6 means two completely different sequences).
Figure X shows what edit distance is required for positive change in TIR for RBS with high and medium TIR.
For high TIR RBS, the minimum distance that is required for increase of TIR is 2, with edit distance between 2 and 5 giving similar results.
For medium TIR RBS, a distance of just 1 is enough to produce a meaningful increase in TIR.
That means that as the TIR of examined RBSs increases, the algorithm should be "encouraged" to explore sequences which are more dissimilar to the current candidates.
This finding could also indicate that Adaptive Laboratory Evolution may not be able to find very strong RBSs - the low rate of natural mutations will be too slow to discover through more dissimilar sequences \cite{Lee2012}.
In other words, because the examined sequence is relatively short (6bp in a wider 20bp context) the time to accumulate 2 or 5 changes in the RBS region might be prohibitively long.
In such cases, a directed process should be strongly encouraged.\\
Another feature noticeable in our library is prevalence of specific patterns.
These include X, Y and Z (Figure X). 
Interestingly, RBSs containing these patterns are not detected as capable of binding the relevant ribosome anti-sequence by binding energy calculations \cite{Mann2017}.
At present, there is no strong evidence why these patterns might by permissive for high TIRs, nevertheless, it is important to note that the used algorithm does not require that knowledge to effectively find them.\\

>>>>

A genetic part library should display a number of characteristics to be deemed useful.
First quality expected from a good library is a wide range of relevant trait values (in our case it is the TIR) so that relevant genetic devices can be appropriately optimised.
Second important feature of a library is the diversity of offered sequences, which ensures that sequences compatible with different assembly methods and plasmids are available. 
Additionally, it is very beneficial for the library to show a number of parts with similar numerical values for relevant trait, but with different sequences, so that a designer would have to settle on suboptimal part due to, for example, assembly incompatibility.
Finally, a good library has to be reliable - if the numbers provided in the library are unreliable designers will be reluctant to use.
Figure X shows how library conforms to these requirements.
In short, due to focusing on exploitation distribution of our TIRs is skewed toward strong ones, but there is still a significant amount of RBSs with medium and low strengths.
High diversity of our sequences is shown by lack of relative lack of bias in position weight matrices for different bins.
Finally, we were able to keep the average standard deviation for each bin within XX\% with 6 biological replicates for each sample, which we find satisfactory for reliability.\\

\begin{figure}[t]
\centering
\subfloat{\label{main:c}\includegraphics[scale=.4]{plots/Hd_Heatmap.png}}\par\medskip
\begin{minipage}{.5\linewidth}
\centering
\subfloat[]{\label{main:a}\includegraphics[scale=.4]{plots/Hd_Heatmap.png}}
\end{minipage}
\begin{minipage}{.5\linewidth}
\centering
\subfloat[]{\label{main:b}\includegraphics[scale=.4]{plots/Hd_Heatmap.png}}
\end{minipage}
\begin{minipage}{.5\linewidth}
\centering
\subfloat[]{\label{main:b}\includegraphics[scale=.4]{plots/Hd_Heatmap.png}}
\end{minipage}
\caption{Characteristics of the sequences.}
\label{fig:characteristics}
\end{figure}

Our library is also showing two characteristics that may be useful in future RBS design efforts.
One of them is the perceived editing distance between two sequences required for meaningful improvement in the TIR. 
We define the editing distance as Hamming distance, that is, how many positions have to be changed to get from one sequence to the other (Hamming distance of 0 means that the sequences are identical and 6 means two completely different sequences).
Figure X shows what edit distance is required for positive change in TIR for RBS with high and medium TIR.
For high TIR RBS, the minimum distance that is required for increase of TIR is 2, with edit distance between 2 and 5 giving similar results.
For medium TIR RBS, a distance of just 1 is enough to produce a meaningful increase in TIR.
That means that as the TIR of examined RBSs increases, the algorithm should be "encouraged" to explore sequences which are more dissimilar to the current candidates.
This finding could also indicate that Adaptive Laboratory Evolution may not be able to find very strong RBSs - the low rate of natural mutations will be too slow to discover through more dissimilar sequences \cite{Lee2012}.
In other words, because the examined sequence is relatively short (6bp in a wider 20bp context) the time to accumulate 2 or 5 changes in the RBS region might be prohibitively long.
In such cases, a directed process should be strongly encouraged.\\
Another feature noticeable in our library is prevalence of specific patterns.
These include X, Y and Z (Figure X). 
Interestingly, RBSs containing these patterns are not detected as capable of binding the relevant ribosome anti-sequence by binding energy calculations \cite{Mann2017}.
At present, there is no strong evidence why these patterns might by permissive for high TIRs, nevertheless, it is important to note that the used algorithm does not require that knowledge to effectively find them.\\
